% -*- coding: utf-8 -*-
% !TEX program = lualatex
\documentclass[oneside]{book}

\input{manual.sty}
\CodeHigh{lite}
\setcounter{chapter}{5}

\begin{document}

\chapter{Tips and Tricks}
\label{chap:tips}

\section{Control Horizontal Alignment}

You can control horizontal alignment of cells in \texttt{tabularray} with
\href{https://www.ctan.org/pkg/ragged2e}{\texttt{ragged2e}} package,
by redefining some of the following commands:

\begin{codehigh}
\RenewDocumentCommand\TblrAlignBoth{}{\justifying}
\RenewDocumentCommand\TblrAlignLeft{}{\RaggedRight}
\RenewDocumentCommand\TblrAlignCenter{}{\Centering}
\RenewDocumentCommand\TblrAlignRight{}{\RaggedLeft}
\end{codehigh}

Please read the documentation of \texttt{ragged2e} package for more details of
their alignment commands.

\section{Use Safe Verbatim Commands}%
\label{sec:fakeverb}

Due to the limitation of TeX, even if you have passed \texttt{verb} option to a
\texttt{tabularray} table, you still could not use some special characters in a
\verb!\verb! command. As an replacement, you may use \verb|\fakeverb| command from \href{https://www.ctan.org/pkg/codehigh}{\texttt{codehigh}} package.

The \verb|\fakeverb| command will remove the backslashes in the following control symbols before
typesetting its content: \fakeverb{\\\\}, \fakeverb{\\\{}, \fakeverb{\\\}}, \fakeverb{\\\#}, \fakeverb{\\\^} and \texttt{\textbackslash\textvisiblespace}, \fakeverb{\\\%}.
Also the argument of \verb|\fakeverb| command need to be enclosed with curly braces.
Therefore it could be safely used inside \verb|tabularray| tables and other LaTeX commands.

Here is an example of using \verb!\fakeverb! commands inside a \verb|tblr| environment
(you don't need \texttt{verb} option to use \verb!\fakeverb!):

\begin{demohigh}
\begin{tblr}{hlines}
  Special & \fakeverb{\abc{}$&^_^uvw 123} \\
  Spacing & \fakeverb{\bfseries\ \#\%}    \\
  Nesting & \fbox{\fakeverb{$\left\\\{A\right.$\#}}
\end{tblr}
\end{demohigh}

In the above example, balanced curly braces and control words (such as \verb!\bfseries!)
need not to be escaped---only several special characters need to be escaped.
Please read the documentation of \texttt{codehigh} package for more details of
\verb|\fakeverb| commands.%
\footnote{By the way, \fakeverb{\EscVerb} command from
\href{https://www.ctan.org/pkg/fvextra}{\texttt{fvextra}} package is similar to
\fakeverb{\fakeverb} command, but with \fakeverb{\EscVerb} you need to escape every control word.}

\section{Acronym Expansion}

When building a table, \texttt{tabularray} has to render each cell twice:
the first time for measuring the size of the cell and the second time for actually typesetting the table.
As a result, acronym management packages that spell out long definitions on their first use and print short definitions thereafter will be impacted by this process.

To enable expected acronym behavior despite this process, \texttt{tabularray} sets the value of a Boolean variable, \verb|\lTblrMeasuringBool|, to indicate whether it is building a cell to measure its size or render the table.

Although there are multiple acronym management packages, we provide two examples here.

The \texttt{acro} package has an \verb|\acswitchoff| macro to bypass marking an acronym's first use.

\begin{codehigh}
\newcommand\myac[1]{\IfBooleanT\lTblrMeasuringBool{\acswitchoff}\ac{#1}}
\end{codehigh}

The \texttt{acronym} package uses starred versions of its macros to bypass marking an acronym's first use.

\begin{codehigh}
\newcommand\myac[1]{%
    \IfBooleanT{\lTblrMeasuringBool}{\ac*{#1}}%
    \IfBooleanF{\lTblrMeasuringBool}{\ac{#1}}}
\end{codehigh}

In both of these examples, the new \verb|\myac| command enables you to check the \texttt{tabularray} measuring status before expanding it within the \texttt{tblr} environment so you can prevent it from being marked as used before it actually appears in your document.

\end{document}
