% -*- coding: utf-8 -*-
% !TEX program = lualatex
\documentclass[oneside]{book}

\input{manual.sty}
\CodeHigh{lite}
\setcounter{chapter}{7}

\begin{document}

\chapter{History and Future}

\section{The Future}

As a policy, \PP{tabularray} can support at most four TeX Live releases with latest updates.
For example, \PP{tabularray} releases published in \the\year{} could be used in TeX Live
\inteval{\year-3}--\the\year{} with latest updates.

To make the upcoming releases more stable, you are very welcome to test the latest package file
in the repository. To test it, you only need to download the following \FF{tabularray.sty}
and put it into the folder of your TeX documents:\newline
\centerline{\url{https://github.com/lvjr/tabularray/raw/main/tabularray.sty}}

\section{The History}

The change log of \PP{tabularray} package will be updated on the wiki page:\newline
\centerline{\url{https://github.com/lvjr/tabularray/wiki/ChangeLog}}

In version 2025A, \CC{\DefTblrTemplate} was deprecated in favor of \CC{\DeclareTblrTemplate}.

In version 2023A, inner key \K{verb} was marked as obsolete,
and it will be removed in the future. But don't worry too much,
because you can stick to old version by \CC{\usepackage{tabularray}[=2023-03-01]}.

In version 2022A, there were several breaking changes:

\begin{itemize}[nosep]
  \item \CC{\multicolumn} command was removed; it is better to use \CC{\SetCell} command.
  \item \CC{\multirow} command was removed; it is better to use \CC{\SetCell} command.
  \item \CC{\firsthline} command was removed; it is better to use \KV{baseline=T} option.
  \item \CC{\lasthline} command was removed; it is better to use \KV{baseline=B} option.
\end{itemize}

For your old documents, you can still rollback to version 2021 by \CC{\usepackage{tabularray}[=v2021]}.

\end{document}
